## Course Write Up for Practical Machine Learning

I am going to use the caret package as discussed in the class to build up my prediction model.I shall also be using randomForest package. I like using the seed value of 1729, Ramanujan's number to ensure the reproducibility of the results.
The data given asks us to train a model to predict the activities of the individual.

```{r echo=FALSE, message=FALSE}
set.seed(1729)
options(warn=-1)
library(Hmisc)
library(caret)
library(randomForest)
library(foreach)
library(doParallel)
```

Reading the data to train the model. Also, we add NAs where the data is invalid.

```{r}
trainSet <- read.csv("data/pml-training.csv", na.strings=c("#DIV/0!") )
questions <- read.csv("data/pml-testing.csv", na.strings=c("#DIV/0!") )
```


```{r}
for(i in c(8:ncol(trainSet)-1)) {trainSet[,i] = as.numeric(as.character(trainSet[,i]))}

for(i in c(8:ncol(questions)-1)) {questions[,i] = as.numeric(as.character(questions[,i]))}
```

Feature Selection from the training data. I did not use imputation, rather selected the features with complete information.

```{r}
vars <- colnames(trainSet[colSums(is.na(trainSet)) == 0])[-(1:7)]
df1 <- trainSet[vars]
vars
```

df1 defines our modelling data. Let's divide our data into training and testing set. Here I use 75-25 division to create a data partition.

```{r}
index <- createDataPartition(y=df1$classe, p=0.75, list=FALSE )
training <- df1[index,]
testing <- df1[-index,]
```

Now we run the code to build a random forest model with n=150 trees. To enhance the speed of the process, we use the doParallel package.

```{r}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

Testing the performance of the model.

```{r}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

## Submission of results
The confusion matrix tells us the accuracy of the model. The accuracy is very high, which gives me the confidence to submit all my answers. I did not try any other model due to lack of time and also the answers that I got from this model were all correct.

Preparing the answers for submission. The code is provided online on the course page.

```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- questions
x <- x[vars[vars!='classe']]
answers <- predict(rf, newdata=x)

answers

pml_write_files(answers)
```